### 2014-11-19 fundamentals/impl.tex
small fixes

Cyclopts (\TODO{cite}), a proof of principle design and implementation of \Cyclus on such a large computer system,
 uses UW-Madison's HTCondor \gls{HTC} infrastructure to perform sensitivity
studies on \Cyclus' resource exchange optimization solvers. To date, it has been
used run over $10^4$ jobs, or (\TODO{compute time}) total compute hours, using
the \Cyclus library via its resource exchange \gls{API}.
The \gls{HTC} infrastructure has also been utilized to run and collect
information from full \Cyclus simulations running in parallel on $10^3$
machines reliably for order $10^5$ simulations.

@@

Interchangeability is accomplished by providing \glspl{API} that define agent-to-agent
interaction and agent-to-environment interaction, primarily through the
\Class{Agent} and \Class{Trader} interfaces. Figure \ref{fig:agent_uml} shows 
the ``inheritance'' structure of the \Class{Agent} class in a \gls{UML} diagram 
format. It shows how the \Class{Agent} interface
provides a notion of parent-child hierarchichal relationship, where parents can
choose to \textit{deploy} child agents and \textit{decommission} child
agents. For the archetype developer, this interface provides enormous power, 
very simply. The \gls{API} provides the \Class{Agent} with helpful functionality for 
interacting with the \Cyclus simulation kernel while abstracting away unnecessary 
detail.

@@

In this way, a researcher can directly compare two different reactor modeling 
implementations (perhaps \Class{DetailedReactor} and \Class{SimpleReactor}) 
simply by exchanging the two corresponding archetypes. That is, two reactor 
archetypes both inheriting from the \Class{Facility} class are indistinguishable 
from a simulation perspective. Two simulations, one using 
\Class{DetailedReactor} and the other using \Class{SimpleReactor} can therefore 
be compared directly to determine the impact of using one or the other model.
This can be done with any agent type, where agents can be ``Regions,'' 
``Institutions,'' or ``Facilities.''

@@

Although concept of the \gls{RIF} hierarchy is provided explicitly, the \Cyclus kernel is
designed to handle hierarchies of arbitrary depth.  It is capable of 
running a simulation without regions and institutions (i.e.  only facilities)
or a simulation having archetypes that are the parents of regions.  By
convention, institutions are mainly responsible for facility deployment, but
any archetype can have access to the \Cyclus agent deployment mechanism.  It is easy
to implement agents that deploy copies of themselves, or that schedule their
own decommissioning.  Because regions, institutions, and facilities are all
agents, they can all be implemented to exercise the full set of features
offered by the \Cyclus kernel.

@@

All operations performed on every resource object (splitting, combining,
decay, etc.) are tracked in detail as they are performed.  This information includes th
agent that newly created resources are created from  
when they are introduced into the simulation.
The parentage of each resource is also tracked. This makes it possible to
follow the history of resources as they are transferred
between agents.  This powerfully impacts the ability to model 
non-proliferation-related simulation goals.

@@

In order to avoid excess decay calculations, when decay 
is invoked, the material checks to see if it contains any nuclides with decay
constants that are significant with respect to the time delta for the current
decay operation.  If none of the decay constants are significant, no decay
calculation is performed and the material remains unchanged.  This error does
not accumulate because the next time the material's decay function is invoked,
the time delta will be larger. \TODO{want more detail on this decay shortcut?}
\TODO{describe decay algorithm and how it might be improved}.

\Cyclus has no notion of ``tracked'' versus ``untracked'' nuclides.  In \Cyclus, the composition of a  material
composition is represented by an arbitrarily large list (potentially thousands) of
nuclides.  Agents are free to treat nuclides present in materials any way they
please - including ignoring them.  It is the responsibility of archetype
developers to choose how to handle potentially full-fidelity compositions.

In large simulations many material objects may be changing frequently.
Material decay can also contribute significantly to such changes.  In order to
help avoid unnecessary runtime performance and database space impacts,
compositions in \Cyclus have some special features, such as immutability.
Immutable compositions allow multiple material objects to all hold
references to the same composition object safely.  When operations are
performed on a material that do not alter the composition, any new resulting
materials hold a reference to the same, original composition object.
Although the decay operation generates a new, separate composition (preserving
immutability), a reference to the new composition is placed in a cache shared
by the new and original compositions. The next time the original
composition is decayed, this cache will be checked for a prior calculation
with the same cumulative decay time.  If one exists, the cached composition is
used.  This cache is shared by all compositions that transitively come from
the same origin.  Composition immutability in concert with decay history
caching help eliminate many redundant decay calculations in addition to
reducing the total number of composition objects.  Since each composition
object is only recorded in the database once, significant space savings also
occurs. \TODO{do we want a diagram helping show this composition caching?}

>>>>

Cyclopts (\TODO{cite}), a proof of principle design and implementation of \Cyclus on such a large computer system,
 uses UW-Madison's HTCondor \gls{HTC} infrastructure to perform sensitivity
studies on \Cyclus' resource exchange optimization solvers. To date, it has been
used to run over $10^4$ jobs, or (\TODO{compute time}) total compute hours, using
the \Cyclus library via its resource exchange \gls{API}.
The \gls{HTC} infrastructure has also been utilized to run and collect
information from full \Cyclus simulations running in parallel on $10^3$
machines reliably for order $10^5$ simulations.

@@

Interchangeability is accomplished by providing \glspl{API} that define canonical agent-to-agent
interaction and agent-to-environment interaction, primarily through the
\Class{Agent} and \Class{Trader} interfaces. Figure \ref{fig:agent_uml} shows 
the ``inheritance'' structure of the \Class{Agent} class in a \gls{UML} diagram 
format. It shows how the \Class{Agent} interface
provides a notion of parent-child hierarchichal relationship, where parents can
choose to \textit{deploy} child agents and \textit{decommission} child
agents. For the archetype developer, this interface provides enormous power, 
very simply. The \gls{API} provides the \Class{Agent} with helpful functionality for 
interacting with the \Cyclus simulation kernel while abstracting away unnecessary 
detail.

@@

In this way, a researcher can directly compare two different reactor modeling 
implementations (perhaps \Class{DetailedReactor} and \Class{SimpleReactor}) 
simply by exchanging the two corresponding archetypes. That is, two reactor 
archetypes both inheriting from the \Class{Facility} class are indistinguishable 
from a simulation perspective. Two simulations, one using 
\Class{DetailedReactor} and the other using \Class{SimpleReactor} can therefore 
be compared directly to determine the whole-simulation impact of using one or the other model.
This can be done with any agent type, where agents can be ``Regions,'' 
``Institutions,'' or ``Facilities.''

@@

Although concept of the \gls{RIF} hierarchy is provided explicitly, the \Cyclus kernel is
designed to handle hierarchies of arbitrary depth.  It is capable of 
running a simulation without regions and institutions (i.e.  only facilities)
or a simulation having archetypes that are the parents of regions.  By
convention, institutions are mainly responsible for facility deployment, but
all archetypes have access to the agent deployment mechanism.  It is easy
to implement agents that deploy copies of themselves, or that schedule their
own decommissioning.  Because regions, institutions, and facilities are all
just agents, they can all be implemented to exercise the full set of features
offered by the \Cyclus kernel.

@@

All operations performed on every resource object (splitting, combining,
decay, etc.) are tracked in detail as they are performed.  This information
includes the agent that created each resource when it was introduced into the
simulation.  The parentage of each resource is also tracked. This makes it
possible to follow the history of resources as they are transferred between
agents.  This powerfully impacts the ability to model
non-proliferation-related simulation goals.

@@

In order to avoid excess decay calculations, when decay 
is invoked, a material checks to see if it contains any nuclides with decay
constants that are significant with respect to the time change since the last
decay operation.  If none of the decay constants are significant, no decay
calculation is performed and the material remains unchanged.  This error does
not accumulate because the next time the material's decay function is invoked,
the time change will be larger. \TODO{want more detail on this decay shortcut?}
\TODO{describe decay algorithm and how it might be improved}.

\Cyclus has no notion of ``tracked'' versus ``untracked'' nuclides.  In \Cyclus, the composition of a material
is represented by an arbitrarily large list (potentially thousands) of
nuclides.  Agents are free to treat nuclides present in materials any way they
please - including ignoring them.  It is the responsibility of archetype
developers to choose how to handle potentially full-fidelity compositions.

In large simulations many material objects may be changing frequently.
Material decay can also contribute significantly to such changes.  In order to
help avoid unnecessary runtime performance and database space impacts,
compositions in \Cyclus have some special features.
Immutability of compositions allows multiple material objects to all hold
references to the same composition object safely.  When operations are
performed on a material that do not alter the composition, any new resulting
materials hold a reference to the same, original composition object.
Although the decay operation generates a new, separate composition (preserving
immutability), a reference to the new composition is placed in a cache shared
by the new and original compositions. The next time the original
composition is decayed, this cache will be checked for a prior calculation
with the same cumulative decay time.  If one exists, the cached composition is
used.  This cache is shared by all compositions that transitively come from
the same origin.  Composition immutability in concert with decay history
caching help eliminate many redundant decay calculations in addition to
reducing the total number of composition objects.  Since each composition
object is only recorded in the database once, significant space savings also
occurs. \TODO{do we want a diagram helping show this composition caching?}
